---
title: "Tables for Legal Rights of Way (200 Feet Buffer)"
author: "Ben Best"
date: "1/7/2019"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true    
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r load}
# packages, reporting----
library(tidyverse)
library(glue)
library(here)
library(knitr)

# paths -----
dir_out <- here("data/200ft")
ter_0_geo       <- here("data/usa_rgn.geojson")
ter_geo         <- file.path(dir_out, "territories.geojson")
lns_geo         <- here("data/lns.geojson")
cables_geo      <- file.path(dir_out, "cables.geojson")
depth_0_tif     <- here("../nrel-uses/data/layers/depth/_depth_epsg4326.tif")
depth_tif       <- file.path(dir_out, "depth.tif")
cables_b_tif    <- file.path(dir_out, "cables_b.tif")
depth_m_tif     <- file.path(dir_out, "depth_m.tif")
depth_mr_tif    <- file.path(dir_out, "depth_mr.tif")
depth_mr_geo    <- file.path(dir_out, "depth_mr.geojson")
cables_td_geo   <- file.path(dir_out, "cables_lns.geojson")
ply_cables_geo  <- file.path(dir_out, "cables_buf.geojson")
tide_max_tif    <- file.path(dir_out, "tide_max.tif")
tide_tif        <- file.path(dir_out, "tide.tif")
tide_geo        <- file.path(dir_out, "tide.geojson")
cables_tide_geo <- file.path(dir_out, "cables_buf200ft_tide.geojson")
wave_0_geo      <- here("data/wave.geojson")
wave_geo        <- file.path(dir_out, "wave.geojson")
wind_0_geo          <- here("data/wind.geojson")
wind_geo            <- file.path(dir_out, "wind.geojson")
cables_tide_geo     <- file.path(dir_out, "cables_buf_tide.geojson")
cables_wave_geo     <- file.path(dir_out, "cables_buf_wave.geojson")
cables_wind_geo     <- file.path(dir_out, "cables_buf_wind.geojson")
energy_depth_3z_csv <- here("data/energy-depth-cables_all.csv")

dir_csv     <- file.path(dir_out, "cables_buffer200ft_tables")
tbl_es1_csv <- file.path(dir_csv, "table_ES-1.csv")
tbl_es2_csv <- file.path(dir_csv, "table_ES-2.csv")
tbl_1_csv   <- file.path(dir_csv, "table_1.csv")
tbl_2_csv   <- file.path(dir_csv, "table_2.csv")
tbl_3_csv   <- file.path(dir_csv, "table_3.csv")
tbl_4_csv   <- file.path(dir_csv, "table_4.csv")
tbl_5_csv   <- file.path(dir_csv, "table_5.csv")
tbl_6_csv   <- file.path(dir_csv, "table_6.csv")

energy_depth_csv    <- file.path(dir_csv, "energy-depth_3z.csv")
territories_csv     <- file.path(dir_csv, "territories.csv")
regions_csv         <- file.path(dir_csv, "regions.csv")
cables_buf_csv      <- file.path(dir_csv, "cables_buf.csv")
cables_buf_tide_csv <- file.path(dir_csv, "cables_buf_tide.csv")
cables_buf_wave_csv <- file.path(dir_csv, "cables_buf_wave.csv")
cables_buf_wind_csv <- file.path(dir_csv, "cables_buf_wind.csv")

tide_depth_cables_3z_csv <- file.path(dir_csv, "tide_depth_cables_3z.csv")
wave_depth_cables_3z_csv <- file.path(dir_csv, "wave_depth_cables_3z.csv")
wind_depth_cables_3z_csv <- file.path(dir_csv, "wind_depth_cables_3z.csv")

# functions, report ----
csv_to_srclink <- function(csv){
  url_gh  <- "https://github.com/ecoquants/nrel-cables/blob/master"
  url_csv <- str_replace(csv, here(), "")
  glue("Source: [{basename(csv)}]({url_gh}{url_csv})")
}
```

```{r analysis, eval=F}
# libraries, + analysis----
library(sf)
library(units)
library(raster)
select <- dplyr::select
devtools::load_all(here("../nrelutils"))

# functions, analysis ----
recode_territory <- function(d){
  d %>% 
    mutate(
      territory = case_when(
        territory %in% c(
          'Puerto Rico','US Virgin Islands') ~ "Atlantic Islands",
        territory %in% c(
          'Guam','Johnston Atoll','N Mariana Islands','Palmyra Atoll','Wake Island') ~ "Pacific Islands",
        TRUE ~ territory))
}

buffer_gcs <- function(x, b, wrap=T, crs=4326){
  # x = lns_cables_td$geometry[[1]]; b = set_units(200, ft); wrap = T

  bb <- st_bbox(x)
  lat_range <- extendrange(bb[c('ymin','ymax')], f=-1/6) # 5.796732 8.983678
  crs_aea <- glue("+proj=aea +lat_1={lat_range[1]} +lat_2={lat_range[2]} +lat_0={mean(lat_range)} 
                  +lon_0={mean(bb[c('xmin','xmax')])} 
                  +x_0=0 +y_0=0 +datum=WGS84 +ellps=WGS84 +units=m +no_defs")
  y <- st_as_sf(tibble(geom=st_sfc(x, crs=crs))) %>% 
    st_transform(crs_aea) %>% 
    st_buffer(b) %>% 
    st_transform(crs)
  
  if (wrap) y <- nrelutils::sf_wrap(y)

  y$geom
}

geo_to_csv <- function(geo, dir_geo, dir_csv){
  # geo = geos[1]
  
  if (!dir.exists(dir_csv)) dir.create(dir_csv)
  path_in  <- glue("{dir_geo}/{geo}")
  path_out <- glue("{dir_csv}/{path_in %>% basename() %>% fs::path_ext_remove()}.csv")
  
  read_sf(path_in) %>%
    st_drop_geometry() %>% 
    write_csv(path_out)
}

# load data ----

# regions
if (!file.exists(ter_geo)){
  ply_ter <- read_sf(ter_0_geo) %>% 
    recode_territory() %>% 
    group_by(territory) %>% 
    summarize(
      area_km2 = sum(area_km2))
  write_sf(ply_ter, ter_geo)
}
ply_ter <- read_sf(ter_geo)

# cables
if (!file.exists(cables_geo)){
  lns_cables <- read_sf(lns_geo) %>% 
    sf_wrap() # x: [141,296]
  write_sf(lns_cables, cables_geo)
}
lns_cables <- read_sf(cables_geo)

# depth
if (!file.exists(depth_tif)){
  r_depth <- raster(depth_wrld_tif) %>%
    crop(cables_sf %>% as("Spatial"))
  r_depth[r_depth < 0] <- NA
  writeRaster(r_depth, depth_tif, overwrite=T)
}
r_depth <- raster(depth_tif)

# buffer cables for depth extraction
if (!file.exists(cables_b_tif)){
  ply_cables_b <- st_buffer(lns_cables, xres(r_depth)) %>% 
    st_cast("POLYGON")
  r_cables_b   <- fasterize::fasterize(ply_cables_b, r_depth)
  writeRaster(r_cables_b, cables_b_tif, overwrite=T)
}
r_cables_b <- raster(cables_b_tif)

# mask depth with cable buffer
if (!file.exists(depth_m_tif)){
  r_depth_m <- mask(r_depth, r_cables_b)
  writeRaster(r_depth_m, depth_m_tif, overwrite=T)
}
r_depth_m <- raster(depth_m_tif)

# reclassify masked depth
if (!file.exists(depth_mr_tif)){
  tbl_reclass_depth <- tribble(
    ~depth_from, ~depth_to, ~depth_beg,
    0, 100, 0,
    100, 200, 100,
    200, 1000, 200,
    1000, Inf, 1000)
  r_depth_mr <- reclassify(r_depth_m, tbl_reclass_depth) # 104.2/60 min
  writeRaster(r_depth_mr, depth_mr_tif, overwrite=T)
}
r_depth_mr <- raster(depth_mr_tif)

# convert to polygon
if (!file.exists(depth_mr_geo)){
  ply_depth_mr <- rasterToPolygons(r_depth_mr, dissolve=T) %>% 
    st_as_sf() %>% 
    select(depth_beg=layer) %>% 
    mutate(
      depth_beg = as.integer(depth_beg))
  write_sf(ply_depth_mr, depth_mr_geo)
}
ply_depth_mr <- read_sf(depth_mr_geo)

# intersect cables with territory and depth ----
if (!file.exists(cables_td_geo)){
  lns_cables_td <- lns_cables %>% 
    # territory
    st_intersection(ply_ter) %>% 
    group_by(territory) %>% 
    summarize(
      n_lines = n()) %>% 
    # depth
    st_intersection(ply_depth_mr) %>% 
    mutate(
      length_km = (st_length(geometry) / 1000) %>% set_units(km),
      depth_lbl = case_when(
      depth_beg ==    0 ~ "0-100",
      depth_beg ==  100 ~ "100-200",
      depth_beg ==  200 ~ "200-1,000",
      depth_beg == 1000 ~ ">1,000",
      TRUE ~ NA_character_)) %>% 
    arrange(territory, depth_beg) %>% 
    select(region=territory, depth_lbl, depth_beg, length_km)

  write_sf(lns_cables_td, cables_td_geo)
}
lns_cables_td <- read_sf(cables_td_geo)

# buffer cables out 200 ft total width, ie 100 ft ea side ----
if (!file.exists(ply_cables_geo)){
  ply_cables <- lns_cables_td %>% 
    mutate(
      geometry = map(geometry, buffer_gcs, set_units(100, ft)) %>% 
        unlist(recursive = F) %>% 
        st_sfc(crs=4326),
      area_km2 = st_area(geometry) %>% set_units(km^2))

  write_sf(ply_cables, ply_cables_geo)
}
ply_cables <- read_sf(ply_cables_geo)

# prep tide energy (points to raster to vector) ----

# tide: points to max raster
if (!file.exists(tide_max_tif)){
  p = read_csv(here("data/tide_rgn-East_pts.csv")) %>%
    bind_rows(
      read_csv(here("data/tide_rgn-West_pts.csv"))) %>% 
    mutate(
      lon = if_else(lon < 0, lon + 360, lon)) %>% 
    select(lon, lat, pwr_wm2)

  r_cables_b2 <- aggregate(r_cables_b, fact=2, fun=max)

  r_max <- rasterize(select(p, lon, lat), r_cables_b2, select(p, pwr_wm2), max)
  
  writeRaster(r_max, tide_max_tif, overwrite=T)
}
r_max <- raster(tide_max_tif)

# tide: reclassify raster
if (!file.exists(tide_tif)){
  tbl_reclass_tide <- tribble(
    ~tide_from, ~tide_to, ~tide_beg,
    0, 500, 0,
    500, 1000, 500,
    1000, 1500, 1000,
    1500, Inf, 1500)
  
  r_tide <- reclassify(r_m, tbl_reclass_tide) # plot(r_mr)
  
  writeRaster(r_tide, tide_max_tif, overwrite=T)
}
r_tide <- raster(tide_max_tif)

# tide: raster to polygon
if (!file.exists(tide_geo)){
  ply_tide <- rasterToPolygons(r_tide, dissolve=T) %>% 
    st_as_sf() %>% 
    mutate(
      tide_beg = as.integer(layer),
      tide_lbl = case_when(
        tide_beg ==    0 ~ "0-500",
        tide_beg ==  500 ~ "500-1,000",
        tide_beg == 1000 ~ "1,000-1,500",
        tide_beg == 1500 ~ ">1,500")) %>% 
    st_intersection(
      ply_ter %>% 
        select(
          region = territory)) %>% 
    select(region, tide_lbl, tide_beg) %>% 
    arrange(region, tide_beg) %>% 
    mutate(
      area_km2 = st_area(geometry) %>% set_units(km^2))
  
  write_sf(ply_tide, tide_geo)
}
ply_tide <- read_sf(tide_geo)

# prep wave & wind energy (polygon) ----

# prep wave
if (!file.exists(wave_geo)){
  ply_wave <- read_sf(wave_0_geo) %>% # x:[171,297]
    mutate(geometry = st_buffer(geometry, dist=0)) %>% 
    recode_territory() %>% 
    rename(
      region   = territory,
      wave_lbl = energy_lbl,
      wave_beg = energy_num) %>% 
    arrange(region, wave_beg)
  write_sf(ply_wave, wave_geo)
}
ply_wave <- read_sf(wave_geo)

# prep wind
if (!file.exists(wind_geo)){
  ply_wind <- read_sf(wind_0_geo) %>% # x:[199,294]
    mutate(
      region = case_when(
        region %in% c(
          'Atlantic','Pacific') ~ glue("{region} Islands"),
        TRUE ~ region)) %>%
    mutate(geometry = st_buffer(geometry, dist=0)) %>%
    mutate(
      wind_lbl = case_when(
        energy_lbl == "<=7" ~ "0-7",
        energy_lbl == "11-12" ~ ">11",
        TRUE ~ energy_lbl),
      wind_beg = case_when(
        energy_num == 7 ~ 0,
        energy_num == 8 ~ 7,
        energy_num == 9 ~ 8,
        energy_num == 10 ~ 9,
        energy_num == 11 ~ 10,
        energy_num == 12 ~ 11,
        TRUE ~ NA_real_)) %>% 
    arrange(region, wind_beg) %>% 
    select(region, wind_lbl, wind_beg, area_km2)
  write_sf(ply_wind, wind_geo)
}
ply_wind <- read_sf(wind_geo)

# intersect cables with energy ----

# tide x cables
if (!file.exists(cables_tide_geo)){
  ply_cables_tide <- ply_cables %>% 
    select(region, depth_lbl, depth_beg) %>% 
    st_intersection(
      ply_tide %>% 
        select(tide_lbl, tide_beg)) %>% 
    mutate(
      area_km2 = st_area(geometry) %>% set_units(km^2))
  write_sf(ply_cables_tide, cables_tide_geo)
}
ply_cables_tide <- read_sf(cables_tide_geo)

# wave x cables
if (!file.exists(cables_wave_geo)){
  ply_cables_wave <- ply_cables %>% 
    select(region, depth_lbl, depth_beg) %>% 
    st_intersection(
      ply_wave %>% 
        select(wave_lbl, wave_beg)) %>% 
    mutate(
      area_km2 = st_area(geometry) %>% set_units(km^2))
  write_sf(ply_cables_wave, cables_wave_geo)
}
ply_cables_wave <- read_sf(cables_wave_geo)

# wind x cables
if (!file.exists(cables_wind_geo)){
ply_cables_wind <- ply_cables %>% 
  select(region, depth_lbl, depth_beg) %>% 
  st_intersection(
    ply_wind %>% 
      select(wind_lbl, wind_beg)) %>% 
  mutate(
    area_km2 = st_area(geometry) %>% set_units(km^2))
  write_sf(ply_cables_wind, cables_wind_geo)
}
ply_cables_wind <- read_sf(cables_wind_geo)

# geojson to csv ----
geos <- list.files(here("data/200ft"), ".*\\.geojson") %>% 
  setdiff(c("cables.geojson", "cables_lns.geojson", "depth_mr.geojson"))
walk(geos, geo_to_csv, dir_geo = dir_out, dir_csv = dir_csv)

# csv tweaks ----
file.copy(energy_depth_3z_csv, energy_depth_csv)
read_csv(territories_csv) %>% 
  rename(region=territory) %>% 
  write_csv(regions_csv)
read_csv(regions_csv)
file.remove(territories_csv)

csvs_in  <- list.files(here("data"), ".*_depth_cables.csv", full.names = T)
csvs_out <- glue("{dir_csv}/{basename(csvs_in) %>% fs::path_ext_remove()}_3z.csv")
file.copy(csvs_in, csvs_out)
```

## ES-1. EEZ & Cable Area

```{r}
options(digits=3)

tbl_es1 <- read_csv(regions_csv) %>% 
  left_join(
    read_csv(cables_buf_csv) %>% 
      group_by(region) %>% 
      summarize(
        cable_km2 = sum(area_km2)),
    by="region") %>% 
  mutate(
    cable_pct = cable_km2 / area_km2 * 100)
write_csv(tbl_es1, tbl_es1_csv)

tbl_es1 %>% 
  kable(
    format.args=list(big.mark=','), 
    caption="Table ES-1: U.S. Exclusive Economic Zone area and cable setbacks area, by region.")
```

`r csv_to_srclink(tbl_es1_csv)`

## ES-2. Energy Resource & Cable Overlap Area

```{r}
eez_km2 <- sum(tbl_es1$area_km2)

tbl_es2 <- read_csv(energy_depth_csv) %>% 
  group_by(energy=form) %>% 
  summarize(viable_km2 = sum(area_km2)) %>% 
  left_join(
    bind_rows(
      read_csv(cables_buf_tide_csv) %>% 
        filter(depth_beg == 0, tide_beg >= 500) %>% 
        summarize(cable_km2 = sum(area_km2)) %>%
        mutate(energy = "tide"),
      read_csv(cables_buf_wave_csv) %>% 
        filter(depth_beg <= 100, wave_beg >= 10) %>% 
        summarize(cable_km2 = sum(area_km2)) %>%
        mutate(energy = "wave"),
      read_csv(cables_buf_wind_csv) %>% 
        filter(depth_beg <= 200, wind_beg >= 7) %>% 
        summarize(cable_km2 = sum(area_km2)) %>%
        mutate(energy = "wind")),
    by="energy") %>%
  mutate(
    viable_pct_eez   = viable_km2 / eez_km2 * 100,
    cable_pct_viable = cable_km2 / viable_km2 * 100) %>% 
  select(
    energy,
    viable_km2, viable_pct_eez,
    cable_km2, cable_pct_viable)

write_csv(tbl_es2, tbl_es2_csv)

tbl_es2 %>% 
  kable(
    format.args=list(big.mark=','), 
    caption="Table ES-2: Definitions of ‘viable’ resource areas for offshore wind, tidal, and wave energy, resource totals, and their cable setback overlap in the U.S. EEZ.")
```

`r csv_to_srclink(tbl_es2_csv)`

## 1. Regions & Cable Lengths

```{r}
tbl_1 <- read_csv(regions_csv) %>% 
  select(
    region, region_km2=area_km2) %>% 
  left_join(
    read_csv(cables_buf_csv) %>% 
      group_by(region) %>% 
      summarize(
        cable_km = sum(length_km)),
    by="region")

write_csv(tbl_1, tbl_1_csv)

tbl_1 %>% 
  kable(
    format.args=list(big.mark=','), 
    caption="Table 1. Summary of U.S. regions data, including EEZ area, cable lengths and availability of resource data.")
```

`r csv_to_srclink(tbl_1_csv)`

## 2. Regions & Cable Overlap Area

```{r}
tbl_2 <- read_csv(regions_csv) %>% 
  select(
    region, region_km2=area_km2) %>% 
  left_join(
    read_csv(cables_buf_csv) %>% 
      group_by(region) %>% 
      summarize(
        cable_km2 = sum(area_km2)),
    by="region") %>% 
  mutate(
    cable_pct_region = cable_km2 / region_km2 * 100)

write_csv(tbl_2, tbl_2_csv)

tbl_2 %>% 
  kable(
    format.args=list(big.mark=','), 
    caption="Table 2. Area of regions and cable setback within the EEZ.")
```

`r csv_to_srclink(tbl_2_csv)`

## 3. Binned Resource & Cable Overlap Area

```{r}
tbl_3 <- read_csv(energy_depth_csv) %>% 
  group_by(form, energy, depth) %>% 
  summarize(viable_km2 = sum(area_km2)) %>% 
  ungroup() %>% 
  mutate(
    energy = recode(energy, `11-12`=">11")) %>% 
  left_join(
    bind_rows(
      read_csv(cables_buf_tide_csv) %>% 
        filter(depth_beg == 0, tide_beg >= 500) %>% 
        group_by(energy=tide_lbl, depth=depth_lbl) %>% 
        summarize(cable_km2 = sum(area_km2)) %>%
        mutate(form = "tide"),
      read_csv(cables_buf_wave_csv) %>% 
        filter(depth_beg <= 100, wave_beg >= 10) %>% 
        group_by(energy=wave_lbl, depth=depth_lbl) %>% 
        summarize(cable_km2 = sum(area_km2)) %>%
        mutate(form = "wave"),
      read_csv(cables_buf_wind_csv) %>% 
        filter(depth_beg <= 200, wind_beg >= 7) %>% 
        group_by(energy=wind_lbl, depth=depth_lbl) %>% 
        summarize(cable_km2 = sum(area_km2)) %>%
        mutate(form = "wind")),
    by=c("form","energy","depth")) %>%
  mutate(
    cable_pct_viable = cable_km2 / viable_km2 * 100) %>% 
  select(
    form, energy, depth,
    viable_km2, 
    cable_km2, cable_pct_viable)

write_csv(tbl_3, tbl_3_csv)

tbl_3 %>% 
  kable(
    format.args=list(big.mark=','), 
    caption="Table 3. Area of energy and depth classes with overlap by cable setback, across all assessed
energy resources.")
```

`r csv_to_srclink(tbl_3_csv)`

## 4. Tidal Resource & Cable Overlap Area by Region

```{r}
tbl_4 <- read_csv(tide_depth_cables_3z_csv) %>% 
  filter(
    territory != "ALL",
    energy_factor != "0-500",
    depth_factor %in% c("0-100")) %>% 
  group_by(region=territory, energy=energy_factor) %>% 
  summarize(viable_km2 = sum(area_km2)) %>% 
  ungroup() %>% 
  left_join(
    read_csv(cables_buf_tide_csv) %>% 
      filter(depth_beg == 0, tide_beg >= 500) %>% 
      group_by(region, energy=tide_lbl) %>% 
      summarize(cable_km2 = sum(area_km2)),
    by=c("region","energy")) %>%
  mutate(
    cable_pct_viable = cable_km2 / viable_km2 * 100) %>% 
  select(
    region, energy,
    viable_km2, 
    cable_km2, cable_pct_viable)

write_csv(tbl_4, tbl_4_csv)

tbl_4 %>% 
  kable(
    format.args=list(big.mark=','), 
    caption="Table 4. Viable tidal resource area and setbacks by region and resource bin.")
```

`r csv_to_srclink(tbl_4_csv)`

## 5. Wave Resource & Cable Overlap Area by Region

```{r}
tbl_5 <- read_csv(wave_depth_cables_3z_csv) %>% 
  filter(
    territory != "ALL",
    energy_factor != "0-10",
    depth_factor %in% c("0-100","100-200")) %>% 
  group_by(region=territory, energy=energy_factor) %>% 
  summarize(viable_km2 = sum(area_km2)) %>% 
  ungroup() %>% 
  left_join(
    read_csv(cables_buf_wave_csv) %>%
      filter(depth_beg <= 100, wave_beg >= 10) %>%
      group_by(region, energy=wave_lbl) %>%
      summarize(cable_km2 = sum(area_km2)),
    by=c("region","energy")) %>%
  mutate(
    cable_pct_viable = cable_km2 / viable_km2 * 100) %>% 
  select(
    region, energy,
    viable_km2, 
    cable_km2, cable_pct_viable)

write_csv(tbl_5, tbl_5_csv)

tbl_5 %>% 
  kable(
    format.args=list(big.mark=','), 
    caption="Table 5. Viable wave resource area and setbacks by region and resource category.")
```

`r csv_to_srclink(tbl_5_csv)`

## 6. Wind Resource & Cable Overlap Area by Region

```{r}
tbl_6 <- read_csv(wind_depth_cables_3z_csv) %>% 
  filter(
    territory != "ALL",
    energy_factor != "<=7",
    depth_factor %in% c("0-100","100-200","200-1,000")) %>% 
  group_by(region=territory, energy=energy_factor) %>% 
  summarize(viable_km2 = sum(area_km2)) %>% 
  ungroup() %>% 
  mutate(
    energy = recode(energy, `11-12`=">11")) %>%
  left_join(
    read_csv(cables_buf_wind_csv) %>%
      filter(depth_beg <= 200, wind_beg >= 7) %>%
      group_by(region, energy=wind_lbl) %>%
      summarize(cable_km2 = sum(area_km2)),
    by=c("region","energy")) %>%
  mutate(
    cable_pct_viable = cable_km2 / viable_km2 * 100) %>% 
  select(
    region, energy,
    viable_km2, 
    cable_km2, cable_pct_viable)

write_csv(tbl_6, tbl_6_csv)

tbl_6 %>% 
  kable(
    format.args=list(big.mark=','), 
    caption="Table 6. Viable wind resource area and setbacks by region and resource category.")
```

`r csv_to_srclink(tbl_6_csv)`

