---
title: "report_extra.Rmd"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# TODO

## High Priority
- get # cables mi/km before/after analysis

## Low Priority
- fix land: US Virgin Islands (Puerto Rico?) and redo maps
- simplify eez+land for PDF maps
- simplify cable polys for PDF maps



# Old TODO References
* [python - Shapely polygons crossing the antimeridian - Geographic Information Systems Stack Exchange](https://gis.stackexchange.com/questions/226605/shapely-polygons-crossing-the-antimeridian)
* [r - Use different center than the prime meridian in plotting a world map - Stack Overflow](https://stackoverflow.com/questions/10620862/use-different-center-than-the-prime-meridian-in-plotting-a-world-map)
* [ogr - How to change shapefile longitude from -180 - 180 to 0 - 360? - Geographic Information Systems Stack Exchange](https://gis.stackexchange.com/questions/79447/how-to-change-shapefile-longitude-from-180-180-to-0-360)
* [sf::st_transform not honoring +lon_wrap · Issue #280 · edzer/sfr](https://github.com/edzer/sfr/issues/280)
* [Moving The Earth (well, Alaska & Hawaii) With R | rud.is](https://rud.is/b/2014/11/16/moving-the-earth-well-alaska-hawaii-with-r/)
* [Administrative Maps and Projections in R - AriLamstein.com](https://www.arilamstein.com/blog/2015/07/23/administrative-level-1-maps-and-projections/)
* [Mapping Data in R · Adam Olson](http://adamolson.org/2015/07/15/post_about_maps/)
* [Mapping All Fifty U.S. States](https://cran.r-project.org/web/packages/fiftystater/vignettes/fiftystater.html)
* [fiftystater/make_mapfile.R at master · wmurphyrd/fiftystater](https://github.com/wmurphyrd/fiftystater/blob/master/data-raw/make_mapfile.R)
* [Mapping All Fifty U.S. States](https://cran.r-project.org/web/packages/fiftystater/vignettes/fiftystater.html)
* [r - Relocating Alaska and Hawaii on thematic map of the USA with ggplot2 - Stack Overflow](https://stackoverflow.com/questions/13757771/relocating-alaska-and-hawaii-on-thematic-map-of-the-usa-with-ggplot2/13767984#13767984)
* [ohiprep/model_create_regions.py at 66bcf4e39b63b154437dda55aec6ed825ac47f05 · OHI-Science/ohiprep](https://github.com/OHI-Science/ohiprep/blob/66bcf4e39b63b154437dda55aec6ed825ac47f05/globalprep/spatial/NCEAS-Regions_v2014/model_create_regions.py)


## Study Area and Submarine Cables

```{r tblDataSources}
read_csv('../data/data_sources.csv') %>%
  kable(
    caption = 'Data sources from preliminary report.') # , booktabs=T)
```

```{r mapStudyArea_leaflet, eval=F, fig.cap='Map of NOAA Charted Submarine cables in the United States as of December 2012.'}
lns_rgn = read_sf(lns_rgn_geo)

pal <- colorFactor(
  colorRampPalette(brewer.pal(11,'Spectral'))(length(usa$territory)),
  usa$territory)

m = leaflet(usa) %>%
  addProviderTiles('Esri.OceanBasemap') %>%
  addPolylines(
    data = lns_rgn, group='cables',
    color='black', opacity = 1, weight=1,
    popup = ~sprintf(
      "scaleBand: %s<br>description: %s<br>effectiveDate: %s",
      scaleBand, description, effectiveDate)) %>%
  addPolygons(
   label = ~sprintf('%s', territory),
   color = ~pal(territory)) #%>%
if (out_html){
  m
} else {
  # [How to save Leaflet in R map as png or jpg file? - Stack Overflow](https://stackoverflow.com/questions/31336898/how-to-save-leaflet-in-r-map-as-png-or-jpg-file)
  m_html = '../tmp/mapStudyArea.html'
  m_png  = '../tmp/mapStudyArea.png'
  saveWidget(m, m_html, selfcontained=F)
  webshot(m_html, file=m_png, cliprect="viewport")
}
```


## Cable Buffer

```{r mapCableBuffs_leaflet, eval=T}

#out_html=F
map_cbls = function(){
  
  if (out_html){
    m = leaflet() %>%
    addProviderTiles("Stamen.TonerLite", group = "B&W") %>%
    addProviderTiles('Esri.OceanBasemap', group = "Ocean") %>%  #
    addPolygons(
      data = dx2, group='Min. Cable (2x)',
      color='red', stroke=F, fillOpacity = 0.5) %>%
    addPolygons(
      data = dx3, group='Rec. Cable (3x)',
      color='red', stroke=F, fillOpacity = 0.5) %>%
    addPolylines(
      data = lns_d1x, group='Cables',
      color='black', opacity = 1, weight=1,
      popup = ~sprintf(
        "scaleBand: %s<br>description: %s<br>effectiveDate: %s<br>Territory: %s<br>Buffer (1x): %d",
        scaleBand, description, effectiveDate, territory, buf1x)) %>%
    addLayersControl(
      baseGroups = c("B&W", "Ocean"),
      overlayGroups = c('Min. Cable (2x)', 'Rec. Cable (3x)', 'Cables'),
      options = layersControlOptions(collapsed = FALSE))
    
    m
  } else {
    
    bb = st_bbox(dx3)    
    p = ggplot() +
      geom_sf(data=wrld2, fill='lightgray') +
      geom_sf(data=dx3, fill='red', color='red') +
      coord_sf(
        xlim=bb[c('xmin','xmax')], 
        ylim=bb[c('ymin','ymax')])
    p
  }
}
map_cbls()
```

## Overlap of Cable Buffer with Renewable Energy

### Wind

TODO:
- [Visualise sf objects — ggsf • ggplot2](http://ggplot2.tidyverse.org/reference/ggsf.html)

- [plotly - stacked bar chart](https://plot.ly/r/bar-charts/#stacked-bar-chart)
- [plotly - horizontal bar chart](https://plot.ly/r/horizontal-bar-charts/#colored-horizontal-bar-chart)
- [plotly - percentages bar chart](https://plot.ly/ggplot2/geom_bar/#precentages)

```{r land selectFeatures}
# resolve land / eez mismatch of borders by filling in eez not high seas
    
# paste(usa_eez$territory, collapse=', ')
#  Alaska, East, Guam, Gulf of Mexico, Hawaii, 
#  Johnston Atoll, N Mariana Islands, Palmyra Atoll, Puerto Rico, 
#  US Virgin Islands, Wake Island, West

ter = 'West' 
land_ter_geo = sprintf('../data/land_%s.geojson', str_replace_all(ter,' ','-'))
#eez_buf_n01 = st_buffer(eez %>% filter(territory==ter), dist = -0.1)
# plot(eez['territory'])
# plot(eez_buf_n01['territory'])

eez_ter = usa_eez %>% filter(territory==ter) # plot(eez_ter['territory'])
bb = st_bbox(eez_ter)
# TODO: st_cast(x=st_bbox, to=st_polygon)
bb_ply = st_sf(
  tibble(
    territory = ter,
    geom = st_sfc(st_polygon(list(rbind(
      c(bb['xmin'], bb['ymin']),
      c(bb['xmax'], bb['ymin']),
      c(bb['xmax'], bb['ymax']),
      c(bb['xmin'], bb['ymax']),
      c(bb['xmin'], bb['ymin'])))), crs = 4326)))

bb_dif = st_difference(bb_ply, eez_ter) %>%
  st_cast('POLYGON', do_split=T)

i_rm = selectFeatures(bb_dif, index=T)
land1 = bb_dif[setdiff(1:nrow(bb_dif), i_rm),] %>%
  group_by(territory) %>%
  summarize()

write_sf(land1, land_ter_geo, delete_dsn=T)
plot(land1)

```

```{r land rm_wrld2}
land1 = read_sf(land_ter_geo)

geos = sprintf(
  '../data/land_%s.geojson', 
  str_replace_all(
    c('Alaska', 'West', 'East', 'Gulf of Mexico'),' ','-'))

ter_sf = function(ter){
  geo = sprintf('../data/land_%s.geojson', str_replace_all(ter, ' ','-'))
  read_sf(geo)
}

ter_geo2shp = function(ter){
  geo = sprintf('../data/land_%s.geojson', str_replace_all(ter, ' ','-'))
  shp = sprintf('../data/land_%s.shp', str_replace_all(ter, ' ','_'))
  read_sf(geo) %>%
    write_sf(shp, delete_dsn=T)
}
ter_geo2shp('Alaska')
ter_geo2shp('West')
ter_geo2shp('East')
ter_geo2shp('Gulf of Mexico')

ter_geo2shp('Johnston Atoll')
ter_geo2shp('N Mariana Islands')
ter_geo2shp('Palmyra Atoll')
ter_geo2shp('Puerto Rico')
ter_geo2shp('US Virgin Islands')
ter_geo2shp('Wake Island')

ter_geo2shp('Guam')
ter_geo2shp('Hawaii')
#  Guam, Hawaii



ter='Alaska'
ter_sf(ter) %>%
  write_sf(
    sprintf('../data/land_%s.shp', str_replace_all(ter,' ','-')), 
    delete_dsn=T)

alaska = ter_sf('Alaska') # plot(alaska)
west = ter_sf('Alaska')
east = ter_sf('East')
gomex = ter_sf('Gulf of Mexico')

nam = rbind(alaska, west, east, gomex)
plot(nam)

wrld2 %>% 
  filter(ID=='USA') %>%
  write_sf('../data/usa2.shp')

wrld2 %>% 
  filter(ID!='USA') %>%
  write_sf('../data/wrld2_notusa2.shp')

nam = st_union(st_combine(nam))
usa2 = st_difference(
  , 
  )
write_sf(usa2, '../data/land_usa2.geojson', delete_dsn=T)

write_sf(wrld2, '../data/land_world2.geojson', delete_dsn=T)

wrld3 = st_difference(wrld2, land1)
mapview(wrld3)
                      
wrld2_i = wrld2 %>% 
  filter(
    ID!='USA',
    st_intersects(wrld2, bb_ply, sparse=F)[,1])



land2 = st_difference(land1, st_union(st_combine(wrld2_i)))

write_sf(land1, land_ter_geo, delete_dsn=T)
plot(land2)
#land1_e = editFeatures(land1)


land4 = read_sf('../data/land_wrld2_usaeez_platecarree.shp') %>% as('Spatial') %>%
  spTransform("+proj=longlat +datum=WGS84 +lon_wrap=180")
plot(land4['ID'])

land5 = readOGR('../data','land_wrld2_usaeez') %>%
  spTransform("+proj=longlat +datum=WGS84 +lon_wrap=180")
plot(land5)

land6 = readOGR('../data','land_wrld2_usaeez_platecarree') %>%
  spTransform("+proj=longlat +datum=WGS84 +lon_wrap=180")
plot(land6)

#ogr2ogr -s_srs EPSG:4326 -t_srs "+proj=longlat +ellps=WGS84 +pm=-180 +datum=WGS84 +no_defs" land_wrld2_usaeez_shifted.shp land_wrld2_usaeez.shp

# [ogr - How to change shapefile longitude from -180 - 180 to 0 - 360? - Geographic Information Systems Stack Exchange](https://gis.stackexchange.com/questions/79447/how-to-change-shapefile-longitude-from-180-180-to-0-360#answer-79448)
ogr2ogr -t_srs "+proj=latlong +pm=-360" tmp.shp land_wrld2_usaeez.shp

tmp = readOGR('../data','tmp') %>%
  spTransform("+proj=longlat +datum=WGS84 +lon_wrap=180")
plot(tmp)

ogr2ogr -skipfailures world_part1.shp land_wrld2_usaeez.shp -clipsrc -180 -90 0 90
ogr2ogr -skipfailures world_part2.shp land_wrld2_usaeez.shp -clipsrc 0 -90 180 90
ogr2ogr world_part1_shifted.shp world_part1.shp -dialect sqlite -sql "SELECT ShiftCoords(geometry,360,0), ID, territory FROM world_part1"
ogr2ogr world_0_360_raw.shp world_part2.shp
ogr2ogr -update -append world_0_360_raw.shp world_part1_shifted.shp -nln world_0_360_raw
ogr2ogr world_0_360.shp world_0_360_raw.shp -dialect sqlite -sql "SELECT ST_Union(Geometry), ID, territory FROM world_0_360_raw GROUP BY ID, territory"

land_w = readOGR('../data','world_0_360') 
land_w = read_sf('../data/world_0_360.shp') 
land_w %>% write_sf(land_usaeez_geo, delete_dsn=T)

plot(land_w['ID'])


land7 = readOGR('../data','land_wrld2_usaeez_shifted') %>%
  spTransform("+proj=longlat +datum=WGS84 +lon_wrap=180") %>%
land7 = land7 %>% st_as_sf()
plot(land7['ID'])

ogr2ogr world_part1.shp world.shp -clipsrc -180 -90 0 90
ogr2ogr world_part2.shp world.shp -clipsrc 0 -90 180 90

%>% st_as_sf() %>%
  write_sf('../data/land_wrld2_usaeez_platecarree.shp')


land2 = land %>% as('Spatial') %>%
  spTransform("+proj=longlat +datum=WGS84 +lon_wrap=180")
land2 = land2 %>% st_as_sf()
plot(land2 %>% filter(ID=='Algeria'))

land3 = cleangeo::clgeo_Clean(land2 %>% as('Spatial'))
plot(land3['ID'])
filter(land3, ID=='Algeria') %>% select(ID) %>% plot()

land2 = land2  
  mutate(
    #geometry   = (geometry + c(360,90)) %% c(360) - c(0,90),
    geom_valid = st_is_valid(geometry)) # Algeria, Mali


land2 %>% filter(!geom_valid)
plot(land2 %>% filter(ID=='Algeria') %>% st_buffer(dist=0))
plot( %>% st_buffer(dist=0))

st_is_valid(land2)
write_sf(land2, '../data/land_wrld2_usaeez.geojson', delete_dsn=T)
    st_set_crs(crs_gcs_w)

read_sf()

wave_sf = wave_sf %>%
    mutate(geometry = (geometry + c(360,90)) %% c(360) - c(0,90)) %>% 
    st_set_crs(crs_gcs_w)
```

### Wave

```{r pltWave, fig.cap='Wave energy (kW/m) and area per US territory with cable overlay (minimum - recommended %).'}
plot_cbls(
  csv   = wave_cbls_csv,
  #title = 'Wave Energy by Area and Power Class per US Territory',
  xlab  = 'Wave energy (kW/m)')

#plot_list = lapply(unique(d_p$territory), plot_wave)
if (out_html){
  #htmltools::tagList(lapply(plot_list, ggplotly))
  # individual: plot_wave('ALL') %>% ggplotly()
  
  #p %>% ggplotly()
  p %>% print()
} else {
  #invisible(lapply(plot_list, print))
  p %>% print()
}
```

```{r mapWave leaflet, eval=F}
# wave read, clean, dissolve internal polys, 
wave = read_sf(wave_geo) %>% 
  as('Spatial') %>% 
  clgeo_Clean() %>%
  st_as_sf() %>%
  filter(
    !is.na(territory)) %>%
  mutate(
    energy = factor(energy_num, labels=unique(energy_lbl), ordered=T)) %>%
  group_by(territory, energy) %>%
  summarise() # mapview(wave, zcol='energy')

pal = colorFactor('Spectral', wave$energy, reverse=T)

leaflet(wave %>% filter(territory=='East')) %>% 
  addProviderTiles('Esri.OceanBasemap') %>%
  addPolygons(
    color="#444444", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.5,
    fillColor = ~pal(energy),
    label = ~sprintf(
      "<strong>%s</strong>: %s kW/m<sup>2</sup>",
      territory, energy) %>% lapply(HTML),
    highlightOptions = highlightOptions(
      color = "white", weight = 2, bringToFront = T)) %>%
  addLegend(pal=pal, values=~energy, title='Wave Energy (kW/m)')
# TODO: fix overlap w/ GoMex vs East and Puerto Rico vs US Virgin Islands
```

### Tidal

units: mean power (W/m^2)

TODO: seperate tif into east and west for display in leaflet

```{r mapTide leaflet, eval=F}
#tide = raster(tide_tif)
#plot(tide_tif)
# TODO: in report.Rmd, leaflet viz tide rasters per territory w/ regex in data/tide_ter-*.tif

tide_ter_tif = '/Users/bbest/github/nrel-cables/data/tide_ter-East.tif'
r = raster(tide_ter_tif)
plot(r)
r_l = projectRasterForLeaflet(r)
plot(r_l)

#pal <- colorNumeric('Spectral', values(r_l), na.color='transparent', reverse=T)
#brks = c(0,500,1000,1500,10753)
pal = colorBin(
  'Spectral', values(r_l), bins =c(0,500,1000,1500,11000), 
  pretty=T, na.color='transparent', reverse=T)

leaflet() %>% 
  addProviderTiles('Esri.OceanBasemap') %>%
  addRasterImage(r_l, colors=pal, opacity=0.8, project=F) %>%
  addLegend(pal=pal, values=values(r_l), title='Tidal Energy (W/m^2)')
```


